\subsection{Grammars}
\subsubsection{LL (2)}
A grammar is LL(2) if it can be parsed by an LL parser that uses 2 tokens of lookahead.

- **Is LL(2)**: 
  - The parser can decide which production to use by looking at the current input symbol and the next input symbol (2 tokens total).
  - No left recursion is allowed.
  - No common prefixes of length 2 or less in the right-hand sides of productions for the same non-terminal.

- **Is not LL(2)**:
  - If it requires more than 2 tokens of lookahead to make parsing decisions.
  - If it contains left recursion.
  - If it has common prefixes of length 2 or less in the right-hand sides of productions for the same non-terminal.


\subsubsection{LR (1)}
A grammar is LR(1) if it can be parsed by an LR parser that uses 1 token of lookahead.

- **Is LR(1)**:
  - The parser can decide which action to take (shift, reduce, or accept) by looking at the current state and the next input symbol.
  - No shift/reduce or reduce/reduce conflicts when constructing the LR(1) parsing table.

- **Is not LR(1)**:
  - If there are any shift/reduce or reduce/reduce conflicts in the LR(1) parsing table.
  - If it requires more than 1 token of lookahead to resolve parsing decisions.


\subsubsection{LR (2)}
A grammar is LR(2) if it can be parsed by an LR parser that uses 2 tokens of lookahead.

- **Is LR(2)**:
  - The parser can decide which action to take by looking at the current state and the next two input symbols.
  - No shift/reduce or reduce/reduce conflicts when constructing the LR(2) parsing table.

- **Is not LR(2)**:
  - If there are any shift/reduce or reduce/reduce conflicts in the LR(2) parsing table.
  - If it requires more than 2 tokens of lookahead to resolve parsing decisions.


\subsubsection{LALR (1)}
LALR(1) (Look-Ahead LR) is a subset of LR(1) grammars that can be parsed with a more compact parser.

- **Is LALR(1)**:
  - It is LR(1).
  - When LR(1) states with the same core are merged, no new conflicts are introduced.

- **Is not LALR(1)**:
  - If it's not LR(1).
  - If merging LR(1) states with the same core introduces new conflicts.


\subsubsection{SLR (1)}
SLR(1) (Simple LR) is the simplest and most restrictive subset of LR(1) grammars.

- **Is SLR(1)**:
  - It is LR(0) (can be parsed with no lookahead).
  - Any remaining conflicts in the LR(0) parsing table can be resolved using the FOLLOW sets of the non-terminals.

- **Is not SLR(1)**:
  - If it's not LR(0).
  - If conflicts remain after considering the FOLLOW sets of non-terminals.





